Forecasting the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd
Revise the Appendix and Measurement

This paper offers valuable analysis for political analysts, campaign strategists, and the public, enhancing the understanding of how polling data can inform electoral outcome predictions. However, recent election cycles have also revealed limitations in polling accuracy due to biases, non-response errors, and variations in poll quality. This study proposes a feasible model that addresses these issues to improve the reliability of election forecasts, and the uncertainty beyond the data has also been acknowledged in this paper.
The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides n evaluation of the SurveyUSA polls and @sec-Appendix2 discusses about the idealized methodology on survey. Besides, model diagnostics and summary are also attached in @sec-model.

@@

The goal of measurement here is to translate individual voter opinions into a reliable estimate of electoral college outcomes. This analysis aims to systematically capture public sentiment through polling data and convert it into a structured forecast of electoral support for each candidate.

- Data Source and Quality Assurance: The dataset is sourced from @fivethirty, a trusted platform recognized for its stringent standards in polling data quality. Only polls meeting high-quality criteria are included to ensure broad representation of likely U.S. voters. Each poll provides essential details, including the pollster’s identity, survey dates, sample size, and methodological transparency. These methods include data collection medium (e.g., phone, online), use of voter files, demographic weighting, and adjustments to create a representative sample. While these standards help capture respondents’ voting intentions accurately, it’s important to note that these criteria rely on historical benchmarks, potentially overlooking recent methodological shifts or emerging biases.
- Temporal Limitations and Data Recency: Polling data captures discrete snapshots of voter sentiment rather than continuous updates. Consequently, the analysis reflects voter intentions only at specific moments, which may not capture rapid opinion shifts, particularly as Election Day approaches. To mitigate this limitation, recency adjustments are applied to account for the time elapsed since each poll was conducted, helping to ensure that the data remains reflective of current voter sentiment.
- Participation and Response Bias: Despite the rigorous methodology, certain biases in survey participation are difficult to avoid. Self-selection bias (where individuals choose whether to participate) and social desirability bias (where respondents may adjust their responses to be more socially acceptable) can affect data accuracy. These biases may introduce discrepancies between expressed opinions and actual voting behavior, potentially influencing forecast reliability.
- Regional Sampling Imbalances: The frequency of polling varies significantly by state, with battleground states receiving more frequent attention than states considered “safe.” This uneven distribution can lead to imbalances in regional representation, which might skew state-level and, by extension, national forecasts. Such discrepancies highlight the challenge of producing a uniformly accurate forecast across all regions.

@@

```{r}
#| echo: false
#| label: tbl-prediction
#| tbl-cap: Prediction Outcomes for the 2024 U.S. Presidential Election by State. Showing the predicted percentage of votes for Donald Trump, the number of electoral votes each state holds, and the expected winning candidate (either Kamala Harris or Donald Trump)
#| warning: false
#| message: false

@@

@tbl-numstate shows the numerical predicted winner of the 2024 US Presidential Election by state. Harris is expected to win the 2024 U.S. Presidential Election, securing 387 electoral votes in total, compared to Trump’s 151 electoral votes. The election outcome hinges on several battleground states, where the vote margins are predicted to be narrow.
```{r}
#| echo: false
#| label: tbl-numstate
#| tbl-cap: Numerical Projected Electoral Vote Count for the 2024 U.S. Presidential Election (Based on Polling Data and Historical Trends). 
#| warning: false
#| message: false

@@

Selecting an appropriate sampling technique can help researchers to obtain reliable research results. In general, sampling techniques fall into two categories: probability (random) sampling and non-probability (non-random) sampling. Probability sampling includes methods such as simple random sampling, stratified random sampling, cluster sampling, systematic sampling, and multi-stage sampling. In contrast, non-probability sampling encompasses techniques like quota sampling, snowball sampling, judgment sampling, and convenience sampling.

SurveyUSA uses a blend of probability and non-probability sampling methods. Some respondents are drawn from non-probability online panels, while others are recruited using probability-based telephone sampling. Responses are weighted based on the latest U.S. Census estimates for age, gender, ethnicity, and region, ensuring alignment with the target population. Questions and answer choices are rotated to reduce order bias, recency effects, and latency effects [@surveyusa_methodology].\

* **Advantages:** The diverse sampling approach not only ensures a broad range of opinions is captured but also complements probability-based sampling, which accurately reflects the overall population. Furthermore, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Additionally, rotating questions and answer choices helps mitigate bias, further improving the reliability of the data. Finally, the use of online surveys offers a cost-effective solution for efficient data collection.\
* **Disadvantages:** Phone-based data collection tends to be time-consuming and can be affected by interviewer effects during telephone interviews. Additionally, challenges like non-response issues, such as busy signals or refusals to participate, can hinder the effectiveness of the data collection process.

@@

Addressing non-response in survey research is essential because it introduces potential bias, particularly non-response bias, which can impact the accuracy and representativeness of the results. As mentioned by @groves2012nonresponse, when certain groups are underrepresented due to non-response, it can skew findings, making them less reflective of the actual population. For example, individuals who with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.

@@

Quota Sampling complements this by setting specific quotas to ensure each subgroup meets minimum representation levels within the sample, based on key demographic categories. While not truly random, quota sampling helps ensure that important demographics are represented even when probability sampling is challenging or costly.

@@

```{r}
#| echo: false
#| label: fig-diagnosis-1
#| tbl-cap: Linear regression assumptions hold.
#| warning: false
#| message: false
#| fig.width: 10
#| fig.height: 8

@@

# Combine metrics into a data frame for display, rounding to 2 decimals
metrics_df <- data.frame(
  Metric = c("Mean Squared Error (MSE)", "Mean Absolute Error (MAE)", "R-squared"),
  Value = round(c(mse, mae, r_squared), 2)
)
kable(metrics_df)
```

## Model Summary

```{r, fig.width=12, fig.height=10}
#| echo: false
#| tbl-cap: Coefficient Estimates with 95% Confidence Intervals for Predictors. Predictors with positive coefficients are shown in blue, while predictors with negative coefficients are shown in red. 
#| warning: false
#| message: false

@@

# Plot the coefficients with confidence intervals, excluding the intercept
ggplot(tidy_model, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(aes(color = estimate > 0), size = 1.2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), color = "grey", height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("Negative Coefficient", "Positive Coefficient"),
    name = ""  # This removes the default "estimate > 0" label
  ) +
  labs(x = "Coefficient Value", y = "Predictors") +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "top", size= 20) +
  guides(color = guide_legend(reverse = TRUE))

@@

## Multicollinearity Check on the Training Data
```{r}
#| echo: false
#| fig-cap: Correlation Matrix of Numeric Grade, Sample Size, Transparency Score and `Days Until Election`
#| warning: false
#| message: false

>>>>

This paper offers valuable analysis for political analysts, campaign strategists, and the public, enhancing the understanding of how polling data can inform electoral outcome predictions. However, recent election cycles have also revealed limitations in polling accuracy due to biases, non-response errors, and variations in poll quality. This study proposes a feasible model that addresses these issues to improve the reliability of election forecasts, and the uncertainty beyond the data has also been acknowledged in this paper.

The paper is organized as follows: @sec-data describes the dataset and the processing steps for model preparation. @sec-modelsetup explains the modeling approach, including predictor selection and the regression model structure. @sec-results presents the primary findings, detailing state-by-state predictions and overall electoral projections. @sec-discussion discusses the key findings, regional voting patterns, polling uncertainty and weakness. The @sec-Appendix1 provides n evaluation of the SurveyUSA polls and @sec-Appendix2 discusses about the idealized methodology on survey. Besides, model diagnostics and summary are also attached in @sec-model.

@@

The goal of measurement is to distill individual voter opinions into a reliable estimate of Electoral College outcomes, transforming raw polling data into a clear forecast of each candidate’s support.

To capture public sentiment, surveys ask questions like, “If the election were held today, who would you vote for?” Responses are summarized by pollsters as a percentage of support for each candidate, creating a structured outcome variable that reflects aggregate support.

- Data Source and Quality Assurance: The dataset is sourced from @fivethirty, a trusted platform recognized for its stringent standards in polling data qualityOnly high-quality polls are included to ensure broad representation of likely U.S. voters. Each poll includes essential details, such as pollster identity, survey dates, sample size, and methodology, all documented in the dataset. While these standards effectively capture respondents' voting intentions, they are based on historical benchmarks that may not fully account for recent methodological shifts or emerging biases. Additionally, the dataset provides a numeric grade, transparency score, and pollster rating, calculated by @fivethirty using their internal criteria and scoring mechanisms. 
- Temporal Limitations and Data Recency: Polling data captures discrete snapshots of voter sentiment rather than continuous updates. Consequently, the analysis reflects voter intentions only at specific moments, which may not capture rapid opinion shifts, particularly as Election Day approaches. To mitigate this limitation, recency adjustments are applied to account for the time elapsed since each poll was conducted, helping to ensure that the data remains reflective of current voter sentiment.
- Participation and Response Bias: Despite the rigorous methodology, certain biases in survey participation are difficult to avoid. Self-selection bias (where individuals choose whether to participate) and social desirability bias (where respondents may adjust their responses to be more socially acceptable) can affect data accuracy. These biases may introduce discrepancies between expressed opinions and actual voting behavior, potentially influencing forecast reliability.
- Regional Sampling Imbalances: The frequency of polling varies significantly by state, with battleground states receiving more frequent attention than states considered “safe.” This uneven distribution can lead to imbalances in regional representation, which might skew state-level and, by extension, national forecasts. Such discrepancies highlight the challenge of producing a uniformly accurate forecast across all regions.

@@

```{r}
#| echo: false
#| label: tbl-prediction
#| tbl-cap: 2024 U.S. Presidential Election Predictions by State. Projected vote percentages for Donald Trump, each state’s electoral votes, and the expected winning candidate (Kamala Harris or Donald Trump).
#| warning: false
#| message: false

@@

@tbl-numstate shows the numerical predicted winner of the 2024 US Presidential Election by state. Harris is expected to win the 2024 U.S. Presidential Election, securing 387 electoral votes in total, compared to Trump’s 151 electoral votes. The election outcome hinges on several battleground states, where the vote margins are predicted to be narrow.
```{r}
#| echo: false
#| label: tbl-numstate
#| tbl-cap: Projected Electoral Vote Count for the 2024 U.S. Presidential Election. With 538 total electoral votes, a candidate needs 270 to win.
#| warning: false
#| message: false

@@

Selecting an appropriate sampling technique can help researchers to obtain reliable research results. In general, sampling techniques fall into two categories: probability (random) sampling and non-probability (non-random) sampling. Probability sampling includes methods such as simple random sampling, stratified random sampling, cluster sampling, systematic sampling, and multi-stage sampling. In contrast, non-probability sampling encompasses techniques like quota sampling, snowball sampling, judgment sampling, and convenience sampling [@fowler2013survey].

SurveyUSA utilizes both probability and non-probability sampling methods, sourcing respondents from non-probability online panels and probability-based telephone sampling [@surveyusa_methodology]. SurveyUSA employs simple random sampling, a method in which each individual in the population has an equal chance of selection, by randomly selecting phone numbers [@noor2022simple]. In contrast, the online panel represents convenience sampling, a non-probability method in which self-selected participants voluntarily join the panel and are readily available for surveys [@fowler2013survey]. Responses are weighted based on the latest U.S. Census estimates for age, gender, ethnicity, and region, ensuring alignment with the target population. Questions and answer choices are rotated to reduce order bias, recency effects, and latency effects [@surveyusa_methodology].\

* **Advantages:** The mixed approach offers several benefits. Simple random sampling ensures representativeness by giving all individuals an equal selection chance, a key feature that ensures representativeness and minimizes selection bias. Randomization helping to offset confounding effects from both known and unknown factors [@noor2022simple]. Convenience sampling through online panels further enhances data collection by making it faster and more cost-effective [@sedgwick2013convenience]. Additionally, reweighting the data according to U.S. Census demographics strengthens the credibility of the results by ensuring demographic accuracy. Rotating questions and answer choices helps mitigate bias, improving data reliability.\
* **Disadvantages:** Simple random sampling can be problematic for heterogeneous or dispersed populations [@noor2022simple]. Using phone-based data collection is time-consuming, as it requires contacting many individuals to reach an adequate response rate. Non-response issues, including busy signals, refusals, or hang-ups, are common in phone surveys. This can create non-response bias, where the views of non-participants differ from those of participants, compromising representativeness. The interviewer effect is another concern as respondents may adjust their answers in the presence of a live interviewer [@noor2022simple]. In addition, convenience sampling also presents challenges, primarily by introducing motivation bias. Participants’ willingness to join may depend on their interest in the survey topic, or dissatisfaction they want to voice, potentially skewing the sample’s representativeness [@sedgwick2013convenience].

@@

Addressing non-response in survey research is essential, as it occurs when selected participants do not participate or complete the survey, leading to non-response bias [@groves2012nonresponse]. This bias can compromise accuracy and representativeness by disrupting the information balance needed for the research question, particularly when certain groups are underrepresented, skewing findings away from the actual population [@sedgwick2013convenience]. For example, individuals who with demanding schedules, lower income or limited access to technology might be less likely to respond to a poll. Meanwhile, people with strong opinions or from specific demographic groups could be overrepresented in the results. As a result of non-response, the findings may misinform decision-makers, leading to conclusions that don’t truly represent the diverse views of the entire population. Addressing non-response is thus essential for producing balanced, reliable survey results that can genuinely inform policies, programs, and public opinion.

@@

Quota Sampling complements this by setting specific quotas to ensure each subgroup meets minimum representation levels within the sample, based on key demographic categories [@moser1952quota]. While not truly random, quota sampling helps ensure that important demographics are represented even when probability sampling is challenging or costly.

@@

```{r}
#| echo: false
#| label: fig-diagnosis-1
#| fig-cap: Linear regression assumptions hold.
#| warning: false
#| message: false
#| fig.width: 10
#| fig.height: 7

@@

# Combine metrics into a data frame for display, rounding to 2 decimals
metrics_df <- data.frame(
  Metric = c("Mean Squared Error (MSE)", "Mean Absolute Error (MAE)", "R-squared"),
  Value = round(c(mse, mae, r_squared), 2)
)
kable(metrics_df)
```
\newpage
## Model Summary

```{r, fig.width=12, fig.height=10}
#| echo: false
#| label: fig-model-summary
#| fig-cap: Coefficient Estimates with 95% Confidence Intervals for Predictors. Predictors with positive coefficients are shown in blue, while predictors with negative coefficients are shown in red. 
#| warning: false
#| message: false

@@

# Plot the coefficients with confidence intervals, excluding the intercept
ggplot(tidy_model, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(aes(color = estimate > 0), size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), color = "grey", height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("Negative Coefficient", "Positive Coefficient"),
    name = ""  # This removes the default "estimate > 0" label
  ) +
  labs(x = "Coefficient Value", y = "Predictors") +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "top", size= 20) +
  guides(color = guide_legend(reverse = TRUE))

@@

## Multicollinearity Check on the Training Data
```{r}
#| echo: false
#| fig-cap: Correlation Matrix of Numeric Grade, Sample Size, Transparency Score and Days Until Election
#| warning: false
#| message: false
